Report on 'How to Revise a Total Preorder' \cite{Booth2011}

### 2020-12-15 paper/paper.tex
feat: seminar table

% TODO write introduction, make WAY SHORTER, inspire by kai
\section{Introduction}
%1.2 Ich sehe hier die Gefahr, dass das zu Ausführlich wird.
%Keine große Diskussion von AGM, Darwiche and Pearl und so weiter.
%Keine Beziehung zur Allen und co herstellen ;)
%Kurz und Prägnant präsentieren, was benötgigt wird um zu Verstehen was
%in dem Primärartikel neu ist.

\begin{itemize}
    \item Introduction of the research area of belief change and major sub-areas according to \cite{Darwiche1997}: nonmonotonic logic, probabilistic reasoning and belief revision and placing of \cite{Booth2011} in the area of belief revision. Belief revision as operator based approach that concerns itself with the definition of constrains (so called postulates) on operators that transitions between held beliefs when new information arrives \cite{Darwiche1997}.
    \item \cite{Alchourron1985} as seminal paper for one-step belief revision on belief sets. Definitions of belief sets and difference between propositional beliefs and conditional beliefs. Adherence to the principle of minimal change for currently held beliefs only; nearly no constrains on conditional beliefs.
    \item \cite{Grove1988} introducing an ordered worlds / spheres approach instead of working on a set of held beliefs. \cite{Booth2004} understanding preorders as partitions into equivalence classes on ordered worlds (and using multiple orders). Iterated belief change as outlook.
    \item \cite{Darwiche1997} discussion of iterated belief revision and the definition and need for epistemic states instead of belief sets for handling conditional beliefs. total preorders as plausibility orderings to represent conditional beliefs.
    \item \cite{Booth2011} discusses a framework for tpo-revision that enriches an epistemic state with a framework to create a new tpo for a revision step on the basis of interval orderings between positive and negative representations of possible worlds. Motivation of tpo-revision as 'context' as introduced in \cite{Booth2007}.
    \item Therefor \cite{Booth2011} as work in iterated belief revision, closer placement of \cite{Booth2011} as non-priotised revision on the belief set level.
    \item Related areas to the framework established \cite{Booth2011} with preference aggregation and relation to research in interval orderings.
\end{itemize}
\textbf{Section Goal:} Overview of the research area and placement of \cite{Booth2011} as work on iterated belief revision. Rough overview of the background for the work in \cite{Booth2011}. Shared understanding of: belief revision (and the difference between one-step- and iterated belief revision), belief sets and epistemic states, conditional beliefs in contrast to propositional beliefs, tpos as plausibility orderings, motivation and definition of tpo-revision as tool for iterated belief revision. Section will need to be discussed with the seminar on \cite{Darwiche1997} and reduced depending on how much is shown in their presentation.

see \cite{Kai2020} for inspiration

% TODO write formal background
\section{Formal background}
\subsection{subscection for each thing, e.g. total preorder}
\subsection{Notation}

%
%1.3.1/2 Das ist denke ich gut so.

%1.3.3-1.3.4 Ich finde gut das du Dinge ausschließt.
%Ich würde sogar vorschlagen, nur 1.3.1 und 1.3.2 auszuarbeiten und auch
%im Vortrag zu präsentieren. Das Paper hat wirklich sehr viel Inhalt.
%Aber die Schwerpunktsetzung sei aber dir überlassen.
%
%

% TODO expand this into subheadings and agenda
\section{Enriched epistemic state and $\leq$-faithful tpos}
\begin{itemize}
    \item Notation basics.
    \item Definition of the DP-AGM postulates and what modifications are made to them for the paper (to always be able to get a belief set from the tpo). Show how a belief set can be extracted from an epistemic state. Introduction of the notation of $\leq$ as tpo, $\ast$ as the revision operator for $\leq$ returning a new tpo $\leq^{\ast}_{\alpha}$. 
    \item Notation of the enriched epistemic state $W^{\pm}$, $\preceq$ and conditions on it as well as it's relation to $\leq$. Visual examples using sticks \cite{Booth2011} and ranks (and mentioning that tpos can be represented as linearly ordered set of ranks) \cite{Booth2006}. Definition of a $\leq$-faithful tpo over $W^{\pm}$.
\end{itemize}
\textbf{Section Goal:} Notation and introducing the concept of a $\leq$-faithful tpo over $W^{\pm}$ and how to visualise it.

% TODO expand this into subheadings and agenda
\section{tpo-revision operators: Revision of $\leq$ with $\preceq$}
\begin{itemize}
    \item Introduction of $\ast$ as revision operator for $\leq$ and definition of $\ast_{\preceq}$ as example with visualisation.
    \item Properties of $\ast_{\preceq}$, their explanation and theorem 1 as axiomatisation of the family of revision operators considered in the paper.
    \item Explanation of sentinential view (and distinction to semantic level), definition of $\leq' \circ \beta $ and how it relates to conditional beliefs, mention of $\leq' \circ \top $ as belief set. Show of properties for the family of revision operators using sentinential view. Explanation of Disj1/Disj2 as properties that are easy to show in the sentinential view but not in the semantic formulation.
\end{itemize}
\textbf{Section Goal:} Definition of $\ast_{\preceq}$ as example revision operator and visualisation of it. Show of axiomatic definition of the family of operators in semantic and sentinential view. Establishing theorem 1.

>>>>

% TODO write introduction, make WAY SHORTER, inspire by kai
\section{Introduction}
\subsection{Research context}
A core process for an agent dealing with uncertain knowledge is updating their worldview when new information becomes available, also called belief change. As an agent can be human or a machine work in belief change has impact on both philosophy and artificial intelligence \cite{Ferme2011}. Different approaches to handling belief change have been discussed with notable approaches including nonmonotonic logic, probabilistic reasoning and belief revision \cite{Darwiche1997}.

In belief revision, belief change is modelled using an operator that produces an updated set of beliefs from the state of an agent and new evidence. Research in this area discusses appropriate formalisations of agent state and constrains on individual or families of operators. Depending on if the suggested operator handles only one new evidence or multiple new evidences the approaches are called one-step or iterated belief revision.

A common tool used to encode plausibility assumptions of an agent about possible worlds are total preorders (shortened as tpos) \cite{Booth2011}. The primary article for this report, 'How to Revise a Total Preorder' discusses how to model the change of these total preorders when new evidence becomes available. Being able to derive a new tpo from additional evidence is an approach to handling multiple revision steps and therefor places the article in the field of iterated belief revision.
%1.2 Ich sehe hier die Gefahr, dass das zu Ausführlich wird.
%Keine große Diskussion von AGM, Darwiche and Pearl und so weiter.
%Keine Beziehung zur Allen und co herstellen ;)
%Kurz und Prägnant präsentieren, was benötgigt wird um zu Verstehen was
%in dem Primärartikel neu ist.
%
%\begin{itemize}
%    \item Introduction of the research area of belief change and major sub-areas according to \cite{Darwiche1997}: nonmonotonic logic, probabilistic reasoning and belief revision and placing of \cite{Booth2011} in the area of belief revision. Belief revision as operator based approach that concerns itself with the definition of constrains (so called postulates) on operators that transitions between held beliefs when new information arrives \cite{Darwiche1997}.
%    \item \cite{Alchourron1985} as seminal paper for one-step belief revision on belief sets. Definitions of belief sets and difference between propositional beliefs and conditional beliefs. Adherence to the principle of minimal change for currently held beliefs only; nearly no constrains on conditional beliefs.
%    \item \cite{Grove1988} introducing an ordered worlds / spheres approach instead of working on a set of held beliefs. \cite{Booth2004} understanding preorders as partitions into equivalence classes on ordered worlds (and using multiple orders). Iterated belief change as outlook.
%    \item \cite{Darwiche1997} discussion of iterated belief revision and the definition and need for epistemic states instead of belief sets for handling conditional beliefs. total preorders as plausibility orderings to represent conditional beliefs.
%    \item \cite{Booth2011} discusses a framework for tpo-revision that enriches an epistemic state with a framework to create a new tpo for a revision step on the basis of interval orderings between positive and negative representations of possible worlds. Motivation of tpo-revision as 'context' as introduced in \cite{Booth2007}.
%    \item Therefor \cite{Booth2011} as work in iterated belief revision, closer placement of \cite{Booth2011} as non-priotised revision on the belief set level.
%    \item Related areas to the framework established \cite{Booth2011} with preference aggregation and relation to research in interval orderings.
%\end{itemize}
%\textbf{Section Goal:} Overview of the research area and placement of \cite{Booth2011} as work on iterated belief revision. Rough overview of the background for the work in \cite{Booth2011}. Shared understanding of: belief revision (and the difference between one-step- and iterated belief revision), belief sets and epistemic states, conditional beliefs in contrast to propositional beliefs, tpos as plausibility orderings, motivation and definition of tpo-revision as tool for iterated belief revision. Section will need to be discussed with the seminar on \cite{Darwiche1997} and reduced depending on how much is shown in their presentation.<
%
%see \cite{Kai2020} for inspiration

% TODO write formal background
\section{Formal background}

\subsection{Propositional Logic} % Aussagenlogik
Notation follows as used in \cite{Booth2011}: A propositional language $L$ generated from finitely many propositional variables, lower case greek letters for formulas in $L$ and $\top$, $\bot$ representing tautology and contradiction respectively. $\vdash$ as classical logical consequence, $\equiv$ classical logical equivalence. $Cn$ is used to denote the deductive closure of a formula or set of formulas in $L$.

$W$ is the set of propositional worlds and given a $\alpha \in L$ the set of worlds that satisfy $\alpha$ is denoted by $[\alpha]$.

For any set of worlds $S \subseteq W$ of worlds, $Th(S)$ is the set of sentences true in all worlds in $S$.

\subsection{Total preorders}
A total preorder is a binary relation $\leq$ (here over the set of propositional worlds $W$) that is total, reflexive and transitive (i.e. for all $x, y, z \in W$ either $x \leq y$ or $y \leq x$, $x \leq x$ and if $x \leq y$ and $y \leq z$ then $x \leq z$).

The symbol $<$ is used to denote the strict part of $\leq$ while $\sim$ represents the symmetric closure of $\leq$ (i.e. $x \sim y$ iff $x \leq y$ and $y \leq x$).

A helpful visualisation for tpos is used in \cite{Booth2006}. It uses the fact that tpos can be represented as a linearly ordered set of ranks. Each rank of a tpo $\leq$ is defined as the equivalence classes modulo the symetric closure of $\leq$: $[[x]]_{\sim} = \{y \vert y \sim x\}$. These equivalence classes are then ordered by the relation $[[x]] \leq [[y]]$ iff $x \leq y$

As an example consider a set of propositional worlds $W = \{ x_{1}, x_{2}, x_{3}, x_{4}\}$ and the a tpo $\leq$ with $x_{1} < x_{2} \sim x_{4} < x_{3}$. The equivalent representation as a linearly ordered set of ranks is shown in \ref{tab:visualising-a-tpo-example}.

\begin{table}[h]
\centering
\begin{tabular}{lll}
$R_{1}$                      & $R_{2}$                                                                   & $R_{3}$                      \\ \hline
\multicolumn{1}{|l|}{$x_{1}$} & \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}$x_{2}$\\ $x_{4}$\end{tabular}} & \multicolumn{1}{l|}{$x_{3}$} \\ \hline
\end{tabular}
\caption{Visualising a tpo as a linearly ordered set of ranks}
\label{tab:visualising-a-tpo-example}
\end{table}

% belief set
% belief revision
% conditional beliefs, evidence/new information

% TODO write chapter
\section{Epistemic states and belief sets}

%why epistemic states? difference between conditional and propositional belief sets
%To understand the requirements imposed by iterated revision we should start by
%recalling the distinction between belief sets and epistemic states. A belief set $ characterizes the set of propositions to which an agent is committed at any given time.
%An epistemic state contains, in addition to @, the entire information needed for coherent reasoning, including, in particular, the very strategy of belief revision which the
%agent wishes to employ at that given time. Any such strategy encodes, and is equivalent to, a set of “conditional beliefs”, that is, beliefs that one is prepared to adopt
%conditioned on any hypothetical evidence [ 2,3,15,16,19,20]. To fully specify behavior under successive observations, one must encode, not merely how beliefs are to be
%revised (this is enough for the first stage only) but also how the revision strategy
%itself is to be modified by each new evidence. This amounts, in turns, to specifying
%which conditional beliefs are to be retained and which ones deleted with each piece of
%evidence.

%Definition of the DP-AGM postulates and what modifications are made to them for the paper (to always be able to get a belief set from the tpo). Show how a belief set can be extracted from an epistemic state. Introduction of the notation of $\leq$ as tpo, $\ast$ as the revision operator for $\leq$ returning a new tpo $\leq^{\ast}_{\alpha}$.

% TODO write chapter
\section{Enriched epistemic states and $\leq$-faithful tpos}
%\begin{itemize}
%    \item Notation of the enriched epistemic state $W^{\pm}$, $\preceq$ and conditions on it as well as it's relation to $\leq$. Visual examples using sticks \cite{Booth2011} and ranks (and mentioning that tpos can be represented as linearly ordered set of ranks) \cite{Booth2006}. Definition of a $\leq$-faithful tpo over $W^{\pm}$.
%\end{itemize}
%\textbf{Section Goal:} Notation and introducing the concept of a $\leq$-faithful tpo over $W^{\pm}$ and how to visualise it.

% TODO write chapter
% booth: "on a semantic level, i.e., in terms of how the ordering of individual worlds x, y undergo change"
\section{tpo-revision operators: Revising $\leq$ with $\preceq$}
%\begin{itemize}
%    \item Introduction of $\ast$ as revision operator for $\leq$ and definition of $\ast_{\preceq}$ as example with visualisation.
\subsection{Example operator}
% shortened example from paper page 232 without SPH teil, show that shortened example satisfies the properties later on. USE EXAMPLE from darwiche -> here to illustrate things with guy killing an alien
%TODO this could be OWN contribution!!?

% TODO write chapter
\section{Properties of tpo-revision operators}
%    \item Properties of $\ast_{\preceq}$, their explanation and theorem 1 as axiomatisation of the family of revision operators considered in the paper.

% TODO write chapter
\section{Sentinential view}
%    \item Explanation of sentinential view (and distinction to semantic level), definition of $\leq' \circ \beta $ and how it relates to conditional beliefs, mention of $\leq' \circ \top $ as belief set. Show of properties for the family of revision operators using sentinential view. Explanation of Disj1/Disj2 as properties that are easy to show in the sentinential view but not in the semantic formulation.
%\end{itemize}
%\textbf{Section Goal:} Definition of $\ast_{\preceq}$ as example revision operator and visualisation of it. Show of axiomatic definition of the family of operators in semantic and sentinential view. Establishing theorem 1.
